Emitter Source Geolocation from Imparted Rotor Blade Modulation

### 2016-10-14 Background.tex
added bib and modified for bistatic doppler calculation.

\section{Literature Review}
%this is the literature review section on what rbm is and current research in the area
The rotating blades of the helicopter modulate signals that hit the blades by reflecting them. The reflected signal, through the doppler effect, become shifted in frequency in relationship to the velocity of the reflection point (refs?). The rotor blade modulation (RBM) has a complex form and is influenced by several object parameters. These parameters range from the positioning of a signal source to the physical dimensions of the blades. The resulting modulated signal has been used in radar applications relating to the classification of any aircraft with a rotor blade, propeller or even jet engines (refs?). The doppler shifted signal has also been shown to have a parasitic effect on GPS anti-jamming algorithms (refs?). The role of RBM is based on the specific doppler shift associated with where a dominant reflection occurs along the length of the of the blade, and the associated elevation angle of the transmitter. The modulated signal will then be used to create a mapping between the amount of doppler present and the distance of a transmitter on the ground.

@@

%describe the situation and the position of the objects
There are three physical objects within this scene. The first is the Rotor blade which is the dark green bar depicted in the figure, next are the transmitter in green $T_x$ and the receiver $R_x$ in blue. The measurements depicted as black bars with denoting start and stop points reference the various distances used in the following equations. The dashed green line shows at which altitude the helicopter is hovering at, and the red dashed line transverses the transmitter and the center of rotation for the blades. $\theta_{El}$ is the elevation angle of the transmitter with respect to the rotor blades and $\theta_r$ denotes the specular reflection of the casted and reflected rays in orange. Lastly the line of sight (LOS) ray shows that the transmitter and receiver are in visual line of sight of each other.

>>>>

\section{Literature Review}
%this is the literature review section on what rbm is and current research in the area
The rotating blades of the helicopter modulate signals that hit the blades by reflecting them. The reflected signal, through the doppler effect, become shifted in frequency in relationship to the velocity of the reflection point \cite{YiminZhang2000}. The rotor blade modulation (RBM) has a complex form and is influenced by several object parameters. These parameters range from the positioning of a signal source to the physical dimensions of the rotor blades.

The resulting modulated signal has been used in several RADAR applications. First to determine the specific characteristics of a noncooperative helicopter using time frequency analysis in \cite{Bullard1991} that defines a specific signature for physical parameters of a helicopter. The ability to detect a hovering helicopter based on the blade flashes caused by reflected RADAR pulses off the blades is examined in \cite{Misiurewicz1997} where they used information the time domain to determine if a helicopter was hovering. The micro-doppler produced by the RADAR reflection in \cite{Thayaparan2007} is used to determine the rotation rate of both the main and tail rotor blades using the wavelet transform. Other types of aircraft produce modulations on a returned RADAR pulse, one type known as Jet Engine Modulation (JEM) is the product of a RADAR pulse reflecting off the jet engine blades \cite{NavalAirSystemsCommand1999}. The information can similarly be used to quantify physical aspects of the aircraft when present in order to perform classification.

The doppler shifted signal caused by RBM is shown in \cite{F.BarbieroF.VincentT.Deloues2014} to effect the performance of common adaptive GPS anti-jamming algorithms located on helicopter platforms. They were able to mitigate the effects by estimating the covariance within a few milliseconds but this comes at an extreme computation overhead. And for iterative based anti-jamming approaches the non-stationary signal characteristics cause degradation as well. The affect of RBM on the BER of communications is examined in \cite{YiminZhang2000} where effects of multi path and doppler cause large bit error rates at small elevation angles and large scattering coefficients described in the model.

The role of RBM in this thesis is based on specific doppler shifts associated with where a dominant reflection occurs along the length of the blade, and the associated elevation angle of the transmitter. The received signal is then used to create a mapping between the amount of doppler present and the position of the transmitter on the ground.

@@

%describe the situation and the position of the objects
There are three physical objects within this scene. The first is the Rotor blade which is the dark green bar depicted in the figure, next are the transmitter in green $Tx$ and the receiver $Rx$ in blue. The measurements depicted as black bars with denoting start and stop points reference the various distances used in the following equations. The dashed green line shows at which altitude the helicopter is hovering at, and the red dashed line transverses the transmitter and the center of rotation for the blades. $\theta_{El}$ is the elevation angle of the transmitter with respect to the rotor blades and $\theta_r$ denotes the specular reflection of the casted and reflected rays in orange. Lastly the line of sight (LOS) ray shows that the transmitter and receiver are in visual line of sight of each other.

### 2016-10-14 Radio_Propagation_Simulation.tex
added bib and modified for bistatic doppler calculation.

\section{Introduction}
%introduce the section and what will be discussed here
The simulation software is broken down into three main components, the radio wave propagation model, the scene model, and the ray tracing engine. The wave propagation model is a collection of equations that describe the physical phenomena that affect the propagation when represented by rays. The model used is based upon classical geometric optics to work with the ray tracing engine. The scene model is comprised of a three dimensional environment in order to define the physical rotor blade surface. The receiver and transmitter will also be defined in the same environment with simulated characteristics to match real life conditions. The ray tracing engine uses recursive execution described in (Whitted, 1979), replacing cameras and lights for receivers and transmitters respectively in a three dimensional environment.

\section{Radio Propagation Model}
%define the math behind the physical laws that are simulated within the simulation
The propagation model defined in this project is based on the following physical phenomena, free-space path loss, reflection, and the doppler effect. They are all formally established models described in full in (Bertoni, 2000; Parsons, 2000; Willis, 1991)(OK). Although several effects are described within the model several were left out to simplify the computation due to the overhead of implementation in three dimensional space and because of the assumptions made about the scene model.

@@

\subsubsection{Diffuse Reflection}
A perfect diffuse reflection, also known as Lambertian reflection, will reflect rays in all directions in a uniform hemisphere (J. Lambert, "Photometria Sive de Mensura et Gradibus Luminus Colorum et Umbrae", 1760). The hemisphere is located at the point of the ray hit, producing rays away from the surface. To approximate a particular type of surface the rays power is attenuated by the surface's bi-directional reflectance distribution function (BRDF) (Suffern, 2007; Pharr, 2010; Glassner, 1995)(OK). There are several ways of calculating the BRDF, but the simplest is the Phong BRDF which will be used in this ray tracing implementation(Phong, 1998)(OK). This method for the BRDF is regularly used in the rendering of computer graphics to simulate the reflection of light off surfaces. Since is was not designed for longer wavelengths used in RF communications an assumption is made that small scale features of the surface, relevant to light waves,  produce a reflectance distribution similar for larger wavelengths.

@@

\subsection{Doppler Effect}
%define doppler effect
The Doppler effect is characterized by (FIX Update to Bistatic Doppler)

@@

%formula for doppler on the rotor blade.
\begin{equation}
	f = f_0 + \frac{2\Delta v}{\lambda} %reradiated after reflection...
	\label{eqn:observedShift}	
\end{equation}

where $f_0$ is the initial transmitted frequency, $lambda$ is the wavelength, and $\Delta v$ is the difference in velocity.

@@

\begin{equation}
	\Delta v = r \omega_r (\vec{d_p} \bullet \vec{d_r})
	\label{eqn:deltav}
\end{equation}

%formula description
where $\Delta v$ is the difference in velocity between the point of reflection along the length of the rotor blade in the direction of the reflection characterized by equation \ref{eqn:deltav}. Where $r$ is the radius along the length of the blade where the reflection occurs. $\omega_r$ is the angular velocity of the rotor. $\vec{d_p}$ is the normalized vector that is perpendicular to the rotor in its direction of travel and $\vec{d_r}$ is the normalized vector that is in the direction of the resulting reflected ray.

@@

\section{Ray Tracing}
%describe the core tracing engine and describe the flow of the program
The ray tracing simulation is conducted in the Monte Carlo fashion acting on the scene and propagation models. This method of raytracing is established for creating realistic images in (Suffern, 2007; Pharr, 2010). The simulation is performed over successive iterations. Each iteration is a discrete element called a frame. Where each frame will be updated to represent the change in the position of the rotor blade.

>>>>

\section{Introduction}
%introduce the section and what will be discussed here
The simulation software is broken down into three main components, the radio wave propagation model, the scene model, and the ray tracing engine. The wave propagation model is a collection of equations that describe the physical phenomena that affect the propagation when represented by rays. The model used is based upon classical geometric optics to work with the ray tracing engine. The scene model is comprised of a three dimensional environment in order to define the physical rotor blade surface. The receiver and transmitter will also be defined in the same environment with simulated characteristics to match real life conditions. The ray tracing engine uses recursive execution described in \cite{Whitted1979}(Whitted, 1979), replacing cameras and lights for receivers and transmitters respectively in a three dimensional environment.

\section{Radio Propagation Model}
%define the math behind the physical laws that are simulated within the simulation
The propagation model defined in this project is based on the following physical phenomena, free-space path loss, reflection, and the doppler effect. They are all formally established models described in full in \cite{Bertoni1999}, \cite{Parsons2000}, and \cite{Willis2005} (Bertoni, 2000; Parsons, 2000; Willis, 1991)(OK). Although several effects are described within the model several were left out to simplify the computation due to the overhead of implementation in three dimensional space and because of the assumptions made about the scene model.

@@

\subsubsection{Diffuse Reflection}
A perfect diffuse reflection, also known as Lambertian reflection, will reflect rays in all directions in a uniform hemisphere (J. Lambert, "Photometria Sive de Mensura et Gradibus Luminus Colorum et Umbrae", 1760). The hemisphere is located at the point of the ray hit, producing rays away from the surface. To approximate a particular type of surface the rays power is attenuated by the surface's bi-directional reflectance distribution function (BRDF) \cite{Suffern2007}, \cite{Pharr2010}, \cite{Glassner1995} (Suffern, 2007; Pharr, 2010; Glassner, 1995)(OK). There are several ways of calculating the BRDF, but the simplest is the Phong BRDF which will be used in this ray tracing implementation \cite{Phong:1998:ICG:280811.280980} (Phong, 1998)(OK). This method for the BRDF is regularly used in the rendering of computer graphics to simulate the reflection of light off surfaces. Since is was not designed for longer wavelengths used in RF communications an assumption is made that small scale features of the surface, relevant to light waves,  produce a reflectance distribution similar for larger wavelengths.

@@

\subsection{Doppler Effect}
%define doppler effect
The Doppler effect is characterized by

@@

%formula for doppler on the rotor blade.
\begin{equation}
	f = f_0 + \frac{v_t + v_r}{\lambda} %reradiated after reflection
	\label{eqn:observedShift}	
\end{equation}

where $f_0$ is the initial transmitted frequency, $lambda$ is the wavelength, $v_t$ is the velocity in the direction of the transmitter, and $v_r$ is the velocity in the direction of the reflection. This is similar to the Bistatic RADAR equation for doppler, but since the reflection is not guaranteed to be toward the receiver we calculate the doppler as it makes contact with the surface and as it is re-radiated according the blades direction of travel.

@@

\begin{equation}
	v_t = r \omega_r (\vec{d_p} \bullet \vec{d_t})
	\label{eqn:v_t}
\end{equation}

%formula description
where $r$ is the radius along the length of the blade where the reflection occurs. $\omega_r$ is the angular velocity of the rotor. $\vec{d_p}$ is the normalized vector that is perpendicular to the rotor in its direction of travel and $\vec{d_t}$ is the normalized vector that is in the direction of the transmitted ray.

\begin{equation}
	v_r = r \omega_r (\vec{d_p} \bullet \vec{d_r})
	\label{eqn:v_r}
\end{equation}

where $\vec{d_r}$ is the normalized vector that is in the direction of the reflected ray.

@@

\section{Ray Tracing}
%describe the core tracing engine and describe the flow of the program
The ray tracing simulation is conducted in the Monte Carlo fashion acting on the scene and propagation models. This method of raytracing is established for creating realistic images in \cite{Suffern2007}, \cite{Pharr2010} (Suffern, 2007; Pharr, 2010). The simulation is performed over successive iterations. Each iteration is a discrete element called a frame. Where each frame will be updated to represent the change in the position of the rotor blade.

### 2016-10-14 Simulation.tex
added bib and modified for bistatic doppler calculation.

Looking back at the geometric models for transmitter azimuth angles of 0\textdegree \space, 90\textdegree \space, 180\textdegree \space, and 270\textdegree \space and their simulated counterparts shown in figure \ref{fig:verify}.

>>>>

Looking back at the geometric models for transmitter azimuth angles of 0\textdegree \space \ref{fig:3D_model_0az_doppler}, 90\textdegree \space \ref{fig:3D_model_90az_doppler}, 180\textdegree \space \ref{fig:3D_model_180az_doppler}, and 270\textdegree \space \ref{fig:3D_model_270az_doppler} and their simulated counterparts shown in figure \ref{fig:verify}. We see that the ray traced results match their projections on the maximum and minimum envelope doppler as transmitter elevation angle changes. The doppler values do not match exactly but the trends are very similar with exception to the 90\textdegree \space result. This discrepancy can be explained in the doppler offset of the other graphs in figure \ref{fig:verify}, where they should start at 0Hz they are shifted up to about 200Hz. This shift overshadows the doppler information when the transmitter is located at 90\textdegree \space and the result is not shifted up because the reflection angle is restricted limiting the maximum available doppler.
